Event History Analyses for psychological time-to-event data: A tutorial in R with examples in Bayesian and frequentist workflows

### 2024-10-16 manuscript/SupplementaryMaterial.Rmd
update supp

cfr paper meemory hazard function

@@

@@

>>>>

The shape of a distribution of waiting times can be described in multiple ways [@luceResponseTimesTheir1991].

After dividing time in discrete, contiguous time bins indexed by t, let RT be a discrete random variable denoting the rank of the time bin in which a particular person's response occurs in a particular experimental condition. Discrete-time EHA focuses on the discrete-time hazard function h(t) = P(RT = t| RT $\geq$ t) and the discrete-time survivor function S(t) = P(RT $>$ t) = [1-h(t)].[1-h(t-1)].[1-h(t-2)]...[1-h(1)], and not on the probability mass function P(t) = P(RT = t) = h(t).S(t-1) nor the cumulative distribution function F(t) = 1-S(t). The discrete-time hazard probability gives you the probability that the event occurs (sometime) in bin t, given that the event has not occurred yet in previous bins. Unlike the discrete-time hazard function, which assesses the unique risk of event occurrencce associated with each time bin, the discrete-time survivor function cumulates the bin-by-bin risks of event *non*occurrence to obtain the probability that the event occurs afer bin t.

For two-choice RT data, the discrete-time hazard function can be extended with the discrete-time conditional accuracy function ca(t) = P(correct | RT = t), which gives you the probability that a response is correct given that it has been emitted in time bin t [@kantowitzInterpretationReactionTime2021; @wickelgrenSpeedaccuracyTradeoffInformation1977; @allisonSurvivalAnalysisUsing2010]. This latter function is also known as the micro-level speed-accuracy tradeoff function.

When time is treated as a continuous variable, let RT be a continous random variable denoting a particular person's response time in a particular experimental condition. Because waiting times can only increase, continuous-time EHA does not focus on the cumulative distribution function F(t) = P(RT $\leq$ t) and its derivative, the probability density function f(t) = F(t)', but on the survivor function S(t) = P(RT $>$ t) and the hazard rate function $\lambda$(t) = f(t)/S(t). The hazard rate function gives you the instantaneous rate of event occurrence at time point t, given that the event has not occurred yet.

@@

(ref:descr-fig3-caption) The logit and cloglog link functions.

```{r plot-link-functions, fig.cap = "(ref:descr-fig3-caption)", out.width='80%'}
knitr::include_graphics("../Tutorial_2_Bayesian/figures/linkfunctions.png")
```

@@

(ref:descr-fig4-caption) Prior distributions on the logit and/or cloglog scales (middle column), and their implications on the probability scale after applying the inverse-logit (or logistic) transformation (left column), and the inverse-cloglog transformation (right column).

```{r plot-priors, fig.cap = "(ref:descr-fig4-caption)", out.width='80%'}
knitr::include_graphics("../Tutorial_2_Bayesian/figures/plot_of_priors.png")
```
